Supermassive black holes and their host spheroids \\ I. Galaxy vivisection

### 2015-10-06 papers/MbhMsph/replytoreferee.txt

-

- The referee is correct in saying that the last point in our Conclusion Section is not a new result and,  
  indeed, we didn't mean to deliver this result as a novelty. The Introductions of Papers I and II present 
  some points about black hole mass scaling relations that are currently under debate. One of these points 
  is the level of intrinsic scatter in the black hole mass - spheroid/galaxy luminosity correlation. Some 
  studies (e.g. Erwin & Gadotti 2012, Beifiori et al. 2012, Kormandy & Ho 2013) claimed that the spheroid 
  luminosity is a better tracer of the black hole mass, but the most recent study by Lasker et al. reported 
  that the spheroid luminosity and the galaxy luminosity are equally good proxies of the black hole mass. 
  Our work solves this controversy by showing that the spheroid luminosity and the galaxy luminosity correlate 
  equally well with the black hole mass when only early-type galaxies are considered (Lasker's sample has only 
  4 spiral galaxies). However, when the galaxy sample contains a reasonable number of spiral galaxies, the 
  correlation with the spheroid luminosity has less intrinsic scatter than that with the galaxy luminosity. 
  To our best knowledge, this has never been shown before in the literature. 
  Additional references have been added to the last point in our Conclusion Section.

- The referee is correct, and we are aware of this. Unlike many papers, we believe that we have not under-
  estimated the uncertainties associated to the bulge best-fit parameters. It is possible that some of the 
  results taken from the literature are "grossly wrong", however the exclusion of extreme outliers would 
  affect only the sigma_3 error bars, i.e. the uncertainties associated to the "grade 3" fits, which constitute 
  only 14% of all fits. The exclusion of extreme outliers would have a very small impact on the sigma_1 and 
  sigma_2 errorbars (we learned this from Savorgnan et al. 2013, where we applied a very similar method). 
  In conclusion, the exclusion of extreme outliers would decrease the uncertainties associated to 14% of 
  all galaxies, with little or no effects on the measurement of the intrinsic scatter.

- We prefer not to reference Burstein et al. because they falsely identified many discs due to poor R^(1/4) 
  decompositions.

- A footnote has been added in Section 4.3 to remind the reader why a slope~1 is consistent with a dry-merging 
  formation scenario, although it simply repeats what already stated in the Introduction.

- A (B-V) galaxy color vs. log(galaxy stellar mass) plot is attached. The early-type galaxies are seen to occupy 
  the region of the red sequence, and the spiral galaxies are located in the region of the blue cloud. 
  The correspondence between galaxy morphological type (early vs. late) and position in the color-mass diagram 
  (red sequence vs. blue cloud) is a very well estabilished result and we don't feel the need of showing this 
  (poor) diagram in the paper.

- We fully agree with the referee on this point, as extensively discussed in Section 4.2.1. We don't think that 
  classical and pseudo-bulges can be distinguished based on their Sersic index. However, this has been common 
  practice in the recent literature, and here we are showing that n<2 bulges are not offset from the correlation
  traced by n>2 bulges. The referee might appreciate that we never refereed to our n<2 bulges merely as pseudo-bulges, 
  but we always specified "claimed by some to be pseudo-bulges". In addition, in the text we have questioned the 
  n<2 criterion used by some to identify pseudo-bulges.

- We never claimed that the break of the black hole - bulge mass correlation is due to the bulge mass. 
  In the Introduction, we anticipated that, for consistency with other known scaling relations (Lsph-sigma and 
  Mbh-sigma), a bent Mbh-Lsph correlation is expected, with the bend occurring around Mbh~10^8. 
  In the Mbh-Lsph and Mbh-Msph diagrams, we identified two different sequences (early vs. late) precisely according 
  to the morphology of galaxies. Indeed, the break is caused by the galaxy morphology. 
  We did not identify pseudo-bulges with n<2 bulges. 
  In Figure 5, the red dashed line marks the correlation for ellipticals+lenticulars (as stated in the caption). 
  The correlation for ellipticals only is not shown in this Figure. The spiral galaxies can be seen not following 
  this correlation. 
  Concerns about the lack of a correlation for spiral galaxies are not correct, and have already been addressed in 
  the last paragraph of Section 4.2. The correlation for spiral galaxies has been measured and quantified with 3 
  different linear regression routines. If there was no correlation for spiral galaxies, the uncertainties associated 
  to the slope and intercept of the linear regression would be extremely large.

### 2015-10-06 papers/MbhMsph/mm.tex
holidays
kry

%\shorttitle{Early and late $M_{\rm BH} - M_{\rm *,sph}$ sequence}
\shorttitle{Red and blue $M_{\rm BH} - M_{\rm *,sph}$ sequence}

@@

%\author{G. A. D. Savorgnan\altaffilmark{1} and A. W. Graham\altaffilmark{1} and A. Marconi and E. Sani\altaffilmark{3} and L.K. Hunt}
%\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}
%\email{gsavorgn@astro.swin.edu.au}

@@

Furthermore, many galaxies host both a classical and a pseudo-bulge (e.g.~\citealt{erwin2003,erwin2015,athanassoula2005,Gadotti2009,
macarthur2009,dosanjosdasilva2013,seidel2015}). 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario,
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
%The Bayesian estimator of \cite{linmixerr} fails in performing an inverse $(X|Y)$ linear regression for the subsample of spiral galaxies. 
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper.

@@

      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$. 

>>>>

\shorttitle{Red and blue $M_{\rm BH} - M_{\rm *,sph}$ sequence}

@@

@@

Furthermore, many galaxies host both a classical and a pseudo-bulge (e.g.~\citealt{erwin2003,erwin2015,
athanassoula2005,gadotti2009,macarthur2009,erwin2010,dosanjosdasilva2013,seidel2015}). 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario\footnote{\bf In dry mergers, the black hole and the bulge grow at the same pace, 
increasing their mass in lock steps.},
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
%The Bayesian estimator of \cite{linmixerr} fails in performing an inverse $(X|Y)$ linear regression for the subsample of spiral galaxies. 
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper.

@@

      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$ {\bf (see also \citealt{erwingadotti2012,beifiori2012})}. 

### 2015-10-06 papers/data_paper/datapaper.tex
kry

Supermassive black holes and their host spheroids\footnote{The term ``spheroid'' refers to either an elliptical galaxy or the spheroidal component 
of a galaxy with a disk. } 

@@

%Modern semi-analytic models use the observed black hole mass scaling relations 
%to constrain the black hole accretion rate .

@@

Measuring the photometric and structural properties of a galaxy's spheroidal component requires the ability to 
separate it from the rest of the galaxy. 

@@

As stressed in Section \ref{sec:int}, our investigation is primarily focused on the central spheroidal components 

@@

We assumed \emph{a priori} that each of our galaxies has a unique spheroidal component, 
that we modeled with a S\'ersic profile. \\

@@

We did not find evidence of bars with exponential profiles (e.g.~\citealt{elmegreenelmegreen1985,gadotti2008,kim2015}), 
but this might be due to the low number statistics of barred galaxies in our sample. }\\

@@

Rings were identified from galaxy images and unsharp masks, and modeled with symmetric Gaussian ring profiles. \\

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/comparison_n.eps} 
\caption{2D measurements of the spheroid S\'ersic indices roughly approximate the 1D equivalent-axis measurements. 
The dashed line displays the 1:1 relation.
{\bf The four most obvious outliers are NGC 821, NGC 3377, NGc 4945, and NGC 5576. 
Their individual cases are discussed in the Appendix. }}
\label{fig:n1d2d}
\end{center}
\end{figure}

@@

  \begin{figure}[h]
  \begin{center}
  \includegraphics[width=\fitfigurewidth]{images/n4945_1Dfit.eps}
  \caption{NGC 4945: 
  An edge-on, dusty spiral galaxy that hosts a Seyfert AGN \citep{lin2011}.
  The light profile has an obvious bump that can be ascribed to the bar.
  This bumb cannot be easily modeled with a Ferrer function {\bf nor with a S\'ersic profile}, 
  thus we exclude the data in the range $100'' \lesssim R_{\rm maj} \sim 400''$ form the fit.
  We also exclude the data within the innermost $6''.4$ due to the contribution from the AGN.
  }
  \end{center}
  \end{figure}

>>>>

Supermassive black holes and their host bulges/spheroids\footnote{\bf By the term ``bulge/spheroid'' 
we mean either an elliptical galaxy or the bulge component of disk galaxies, 
with no attempt of distinguishing between classical and disk-like pseudo-bulges. } 

@@

@@

Measuring the photometric and structural properties of a galaxy's spheroidal component 
requires the ability to separate it from the rest of the galaxy. 

@@

{\bf As stressed in Section \ref{sec:int}, our investigation is primarily focused on the central bulge/spheroidal components 

@@

The objects in our sample are either early-type galaxies (elliptical+lenticular), 
or ``early-type spiral'' galaxies 
(i.e.~the morphological classification of our spiral galaxies is within \emph{Sa-Sc}, 
with the only exception of NGC 4945 which is classified as \emph{Scd}), 
therefore -- by definition -- they all have a bulge/spheroidal component, 
unlike ``late-type spiral'' galaxies that can be bulge-less (e.g.~NGC 300).   
We modeled spheroids/bulges with a S\'ersic profile, without attempting to distinguish between classical and pseudo-bulges. } \\

@@

Our galaxy sample lacks of ``late-type spiral galaxies''.
As expected, we did not find evidence of bars with exponential profiles (e.g.~\citealt{elmegreenelmegreen1985,gadotti2008,kim2015}), 
which are typical of late-type spiral galaxies \citep{gadottidesouza2006}, 
and we only modeled bars with flat profiles, which are common features of ``early-type spiral'' galaxies \citep{gadottidesouza2006}. }\\

@@

Rings were identified from galaxy images and unsharp masks, and modeled with symmetric Gaussian ring profiles 
(e.g.~\citealt{sheth2010,kim2014}). \\

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/comparison_n.eps} 
\caption{2D measurements of the spheroid S\'ersic indices roughly approximate the 1D equivalent-axis measurements. 
The dashed line displays the 1:1 relation.
{\bf The four most obvious outliers are NGC 821, NGC 3377, NGC 4945, and NGC 5576. 
Their individual cases are discussed in the Appendix. }}
\label{fig:n1d2d}
\end{center}
\end{figure}

@@

  \begin{figure}[h]
  \begin{center}
  \includegraphics[width=\fitfigurewidth]{images/n4945_1Dfit.eps}
  \caption{NGC 4945: 
  An edge-on, dusty spiral galaxy that hosts a Seyfert AGN \citep{lin2011}.
  The light profile has an obvious bump that can be ascribed to the bar.
  This bumb cannot be easily modeled with a Ferrer function {\bf nor with a S\'ersic profile}, 
  thus we exclude the data in the range $100'' \lesssim R_{\rm maj} \sim 400''$ from the fit.
  We also exclude the data within the innermost $6''.4$ due to the contribution from the AGN.
  }
  \end{center}
  \end{figure}
